---
title: "tissue_specificity_ade"
author: "Lizzy Wilk"
date: "1/30/2023"
output: html_document
---

## Purpose  
Investigate tissue-specific and pre-cystic/cystid PKD kidney expression of the drug targets of sirolimus (aka rapamycin).  
load libraries
```{r}
library(readxl)
library(tidyverse)
library(ComplexHeatmap)
library(viridis)
library(ggplot2)
library(ggalluvial)
library(signatureSearch)
library(signatureSearchData)
```

#### Sirolimus Targets
Signaturesearch's function __get_targets()__ returns for a set of query drug names/ids the corresponding target gene/protein ids.
The required drug-target annotations are from DrugBank, CLUE and STITCH. An SQLite database
storing these drug-target interactions based on the above three annotation resources is available in
the signatureSearchData package.  
’DrugBank’, ’CLUE’, STITCH’
get targets for sirolimus
```{r}
sirol_ss <- signatureSearch::get_targets("sirolimus", database = "all", verbose = TRUE, output = "vector")

```

read in pharos's results for sirlimus / rapamycin (tclin and tchem) and signor's signaling
```{r}
sirol_pharos <- read_csv("../data/rapamycin_pharos/queryresults.csv")

sirol_signor <- read.delim("../data/sirolimus_signor.tsv")
```

combine the unique targets across all data sources
```{r}
target_aggreg <- unique(c(sirol_ss, sirol_pharos$Symbol, sirol_signor$ENTITYB))

```


### Tissue Specific Drug Targets
gtex_tissue_specificity
read in gtex
```{r}
gtex <- read.csv("../data/gtex_tissue_specificity.txt", sep = "\t")

gtex <- gtex[3:nrow(gtex),] 
gtex <- dplyr::select(gtex, !c(2,3))
```


filter gtex tissue specificity just for genes that are drug targets
```{r}
tis_spec_targets <- dplyr::filter(gtex, `X.` %in% target_aggreg)
temp <- tibble::column_to_rownames(tis_spec_targets, var = "X.") 
tis_spec_targets <- lapply(temp, as.numeric) %>% as.data.frame()
row.names(tis_spec_targets) <- row.names(temp)
```

drug targets that have a specificity in more than 1 tissue, and tisues with no specificity (except kidney!)
```{r}
#test <- colVars(as.matrix(tis_spec_targets), useNames = TRUE)
#sort(test)

#focus on tissues with > 1 biased target expression
av <- apply(tis_spec_targets, 2, abs)
#find out how many target biased exp per tissue
t <- apply(av, 2, sum)
#tissues with > 1 biased target expression
keep <- t[t>1]

#select tissues
tis_spec_targets <- dplyr::select(tis_spec_targets, c("kidney", names(keep))) #include kidney even if it has less than 2 biased targets!

```

annotate targets by sirolimus's effect (from signor data)
```{r}
sirolimus_effects <- #sirol_signor %>% dplyr::filter(ENTITYB %in% target_aggreg) %>%
  tis_spec_targets %>% rownames_to_column( var = "ENTITYB") %>% dplyr::left_join(., sirol_signor, by = "ENTITYB") %>%
  dplyr::mutate(Sirolimus_Effect = stringr::str_sub(EFFECT, start = 1, end = 14)) %>% 
  dplyr::select(ENTITYB, Sirolimus_Effect, Effect_Mechanism = MECHANISM) %>% distinct() %>%  mutate_all(list(~na_if(.,""))) %>% column_to_rownames(var = "ENTITYB")
  
row_ha = rowAnnotation(Sirolimus_Effect = sirolimus_effects$Sirolimus_Effect, 
                       Effect_Mechanism = sirolimus_effects$Effect_Mechanism,
                       #colors for each annotation
    col = list(Sirolimus_Effect = c("up-regulates" = rocket(4)[4], "down-regulates" = rocket(4)[1]),
               Effect_Mechanism = c("transcriptional regulation" = rocket(4)[2], "chemical inhibition" =  rocket(4)[3])),
    #legend font sizes
    annotation_legend_param = list(Sirolimus_Effect = list(title = "Sirolimus Effect", title_gp = gpar(fontsize = 14, fontface = "bold"), labels_gp = gpar(fontsize = 12, fontface = "bold")), 
                                   Effect_Mechanism = list(title = "Effect Mechanism", title_gp = gpar(fontsize = 14, fontface = "bold"), labels_gp = gpar(fontsize = 12, fontface = "bold"))),
    #annotation labels
    annotation_name_gp= gpar(fontsize = 12, fontface = "bold")
    )

```


##### Plot heatmap
```{r}
png("../figures/gtex_tissuespecificDEG_sirolimus_heatmap.png", width = 25, height = 40, units = "cm", res = 300)

set.seed(1) #seed for clustering
ComplexHeatmap::Heatmap(tis_spec_targets, row_labels = rownames(tis_spec_targets), right_annotation = row_ha, col = cividis(3),
                        #legend font sizes
                        heatmap_legend_param = list(title_gp = gpar(fontsize = 14, fontface = "bold"), labels_gp = gpar(fontsize = 12, fontface = "bold")),
      #labels/titles
name = "Tissue Specificity", row_title = "Sirolimus Targets", row_title_gp = gpar(fontsize = 16.2, fontface = "bold"),
        column_title = "Tissue Specific Expression of Sirolimus Drug Targets", column_title_gp = gpar(fontsize = 19, fontface = "bold"), column_names_rot = 45,
  #resize other fonts
  row_names_gp = gpar(fontsize = 14, fontface = "bold"),column_names_gp = gpar(fontsize = 15, fontface = "bold"))

dev.off()
```

#### Versions  
```{r}
R.Version()
```


```{r}
installed.packages()[names(sessionInfo()$otherPkgs), "Version"]
```
